---
title: "RCTD_anno"
author: "Jin Huhao"
date: "2024-04-04"
output:
  pdf_document: default
  html_document: default
---

```{r}
# libraries
library(SeuratDisk)
library(Seurat)
library(spacexr)
library(Matrix)
library(dplyr)
library(ggplot2)
```

```{r}
path <- "D:/Desktop/MGI/CODE/"

# Data
data <- LoadH5Seurat(paste0(path, "Data/SS200000116BR_E6.bin200.h5seurat"))
head(data@meta.data, 5)
```

```{r}
# coords: A dataframe or matrix. 
# Rownames are barcodes/pixel names, and there should be 2 columns for x and y.
coords <- data@meta.data %>% select(x, y)
coords
```


```{r}
# counts: A matrix or dgCmatrix representing Digital Gene Expression.
# Rownames should be genes and colnames represent barcodes/pixel names.
counts = GetAssayData(data)
print(head(colnames(counts)))
print(head(rownames(counts)))
print(head(counts))
```
```{r}
# nUMI (optional)
# a named (by pixel barcode) list of total counts at each pixel 
nUMI <- colSums(counts)
head(nUMI) # In this case, total counts per pixel is nUMI
```

```{r}
# SpatialRNA: constructor of SpatialRNA object
# SpatialRNA(coords, counts, nUMI = NULL, use_fake_coords = FALSE, require_int = TRUE)
puck <- SpatialRNA(coords, counts, nUMI)
puck
```

```{r}
print(dim(puck@counts))
hist(log(puck@nUMI,2))
```

```{r}
# load reference (class: SummarizedExperiment)
ref <- readRDS(paste0(path, "Data/5patients_nG3000_PC30_Res0.4_Celltype.rds"))
```

```{r}
metadata_ref <- ref@meta.data
metadata_ref
```

```{r}
cell_types <- metadata_ref %>% select(celltype.singResource)
cell_types

names <- rownames(cell_types)
head(names,5)
```

```{r}
# counts: A matrix or dgCmatrix representing Digital Gene Expression.
# Rownames should be genes and colnames represent barcodes/cell names.
counts_ref <-  ref[["RNA"]]@counts
print(head(colnames(counts_ref)))
print(head(rownames(counts_ref)))
print(head(counts_ref))
```

```{r}
# cell type: A named (by cell barcode) factor of cell type for each cell.

# level: the "level" of the factor would be the possible cell type identities.

# levels(cell_types) contains a cell type 
# with name containing prohibited character /.
# There are T/NK and Tumor/Cholang contains "/"

cell_types$celltype.singResource <- gsub("/", "-", cell_types$celltype.singResource)

levels <- unique(cell_types$celltype.singResource)

cell_types_ref <- factor(cell_types$celltype.singResource, levels = levels)
names(cell_types_ref) <- names

head(cell_types_ref, 20)
```


```{r}
# nUMI (optional)
# a named (by cell barcode) list of total counts at each pixel 

nUMI_ref <- colSums(counts_ref)
head(nUMI_ref)
```

```{r}
# Reference: constructor of Reference object.
# Reference(counts, cell_types, nUMI = NULL, require_int = TRUE, n_max_cells = 10000, min_UMI = 100)
reference <- Reference(counts_ref, cell_types_ref, nUMI_ref)
```
So far, the Reference and SpatialRNA object are successfully constructed!

Next, Create RCTD object and run RCTD

```{r}
# create RCTD
RCTD <- create.RCTD(spatialRNA = puck, reference = reference, max_cores = 1)
```

```{r}
# run RCTD
RCTD <- run.RCTD(RCTD = RCTD, doublet_mode = "doublet")
```
Get the results of RCTD

RCTD has 3 modes:

1. doublet mode:
assign 1-2 cell types per spot
recommended for technologies with high spatial resolution

2. full mode:
assign any number of cell types per spot
recommended for technologies with poor spatial resolution

3. multi mode:
an extension of doublet model, can discover more than 2 cell types per spot 
as an alternative option to full model 

```{r}
saveRDS(RCTD, file = "RData/RCTD.rds")
```

```{r}
RCTD <- readRDS(paste0(path,"RData/RCTD.rds"))
```

```{r}
results <- RCTD@results
norm_weights = normalize_weights(results$weights) 
cell_type_names <- RCTD@cell_type_info$info[[2]] #list of cell type names
spatialRNA <- RCTD@spatialRNA
```

```{r}
RCTD_result_dir <- paste0(path, 'Figure/RCTD')
dir.create(RCTD_result_dir)
```

```{r}
# make the plots 

# cell_type_weights_unthreshold.pdf
# Plots the confident weights for each cell type as in full_mode
plot_weights(cell_type_names, spatialRNA, RCTD_result_dir, norm_weights) 

# cell_type_weights.pdf
# Plots all weights for each cell type as in full_mode. 
plot_weights_unthreshold(cell_type_names, spatialRNA, RCTD_result_dir, norm_weights) 

# cell_type_weights_doublets.pdf
# Plots the weights for each cell type as in doublet_mode.
plot_weights_doublet(cell_type_names, spatialRNA, RCTD_result_dir, 
                     results$weights_doublet, results$results_df) 

# cell_type_occur.pdf
# Plots the number of confident pixels of each cell type in 'full_mode'. 
plot_cond_occur(cell_type_names, RCTD_result_dir, norm_weights, spatialRNA)
```

```{r}
# all_cell_types.pdf
# makes a map of all cell types
plot_all_cell_types(results$results_df, spatialRNA@coords, cell_type_names, RCTD_result_dir) 
```

```{r}
# doublets
# obtain a dataframe of only doublets
doublets <- results$results_df[results$results_df$spot_class == "doublet_certain",] 

# all_doublets.pdf
# Plots all doublets in space
plot_doublets(spatialRNA, doublets, RCTD_result_dir, cell_type_names) 
```

```{r}
# all_doublets_type.pdf
# Plots all doublets in space for each cell type
plot_doublets_type(spatialRNA, doublets, RCTD_result_dir, cell_type_names) 
# a table of frequency of doublet pairs 
doub_occur <- table(doublets$second_type, doublets$first_type) 

# doublet_stacked_bar.pdf
# Plots a stacked bar plot of doublet ocurrences
plot_doub_occur_stack(doub_occur, RCTD_result_dir, cell_type_names) 
```
```{r}
# the probabilities of being first type and second type
probs <- results$weights_doublet
colnames(probs) <- c("first_prob", "second_prob")
head(probs)
```
```{r}
# contains first_type, second_type, first_prob, second_prob and spot_class
# spot_class
res <- cbind(results$results_df %>% select(spot_class, first_type, second_type), probs)
head(res)
```

```{r}
res$final_type <- ifelse(res$first_prob > res$second_prob, as.character(res$first_type), as.character(res$second_type))
head(res)
```

```{r}
# the proportion of first_prod > second_prob
sum(res$first_prob > res$second_prob)/nrow(res)

sum(res$first_type == res$second_type)
```


```{r}
coords <- spatialRNA@coords
print(head(coords))

print(dim(res)[1])
print(dim(coords)[1])
```

```{r}
res$x <- coords$x
res$y <- coords$y
head(res)
```

```{r}
plot_first_type <- 
  ggplot(res, aes(x = x, y = y, color = factor(first_type)))+
  geom_point() +
  theme_minimal()
print(plot_first_type)
```

```{r}
plot_second_type <- 
  ggplot(res, aes(x = x, y = y, color = factor(second_type)))+
  geom_point() +
  theme_minimal()
print(plot_second_type)
```

```{r}
plot_final_type <- 
  ggplot(res, aes(x = x, y = y, color = factor(final_type)))+
  geom_point() +
  theme_minimal()
print(plot_final_type)
```

```{r}
write.csv(res, paste0(path, "Output/RCTD_results.csv"))
```

