---
title: "SingleR_annotation"
author: "Jin Huhao"
date: "2024-04-02"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
### libraries
library(Seurat)
library(dplyr)
library(patchwork)
library(ggplot2)
library(SeuratDisk)
library(SingleR)
library(celldex)
library(pheatmap)
library(SummarizedExperiment)
library(scater)
```

```{r}
# Data
data <- LoadH5Seurat("Data/SS200000116BR_E6.bin200.h5seurat")
# data is already Seurat object
head(data@meta.data, 5)
```
```{r pressure, echo=FALSE}
nCount_RNA <-  colSums(GetAssayData(data)) # 计算每个细胞的RNA读数计数
nFeature_RNA <-  colSums(GetAssayData(data) > 0) # 计算每个细胞检测到的唯一基因数
data@meta.data$nCount_RNA <- nCount_RNA
data@meta.data$nFeature_RNA <- nFeature_RNA
head(data@meta.data, 5)
```

```{r}
# MT
data[["percent.mt"]] <- PercentageFeatureSet(data, pattern = "^MT-")
head(data@meta.data, 5) 
```

```{r}
# save data
saveRDS(data, "RData/E6_bin200.rds")
```

```{r}
# load data
data <- readRDS("RData/E6_bin200.rds")
```

```{r}
VlnPlot(data, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

```{r}
# filter (based on the previous Vlnplot)
data <- subset(data, subset = nFeature_RNA >= 2500 & nFeature_RNA <= 10000 & percent.mt <= 7 & percent.mt >= 2)
VlnPlot(data, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

```{r}
# Normalize
data <- NormalizeData(data, normalization.method = "LogNormalize", scale.factor = 10000)
```

```{r}
# 识别高变基因
data <- FindVariableFeatures(data, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes 
top10 <- head(VariableFeatures(data), 10)
top10
```

```{r}
# plot variable features with and without labels
plot1 <- VariableFeaturePlot(data)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1
plot2
```
```{r}
# scale
all.genes <- rownames(data)
data <- ScaleData(data)
```

```{r}
# Perform linear dimensional reduction
data <- RunPCA(data, features = VariableFeatures(object = data))

# Examine and visualize PCA results a few different ways
print(data[["pca"]], dims = 1:5, nfeatures = 5)
```

```{r}
VizDimLoadings(data, dims = 1:2, reduction = "pca")
```
```{r}
DimPlot(data, reduction = "pca") + NoLegend()
```

```{r}
DimHeatmap(data, dims = 1, cells = 500, balanced = TRUE)
DimHeatmap(data, dims = 1:15, cells = 500, balanced = TRUE)
```

```{r}
# Elbow
# In this example, we might have been justified in choosing anything between PC 3-6 as a cutoff.
ElbowPlot(data)
```
```{r}
data <- FindNeighbors(data, dims = 1:10)
data <- FindClusters(data, resolution = 0.5)
head(Idents(data), 5)
```
```{r}
# UMAP based on the cluster results
data <- RunUMAP(data, dims = 1:10)
```

```{r}
# note that you can set `label = TRUE` or use the LabelClusters function 
# to help label individual clusters
DimPlot(data, reduction = "umap")

VlnPlot(data, features = c("AL391095.2", "IGHM"))
```
```{r}
### Save R data
saveRDS(data, file = "RData/E6_bin200_seurat_result.rds")
```

Using SingleR to annotate cell type

```{r}
# reference annotation data
load("Data/HumanPrimaryCellAtlas_hpca.se_human.RData")
load("Data/BlueprintEncode_bpe.se_human.RData")
```

```{r}
data <- readRDS("RData/E6_bin200_seurat_result.rds")
data
```

```{r}
meta <- data@meta.data
head(meta)

write.csv(meta, "Output/seurat_cluster.csv")
```

```{r}
# UMAP
DimPlot(data, reduction = "umap", label = TRUE)
```
```{r}
# hpca: Human Primary Cell Atlas
data_for_SingleR <- GetAssayData(data, slot="data")
data.hesc <- SingleR(test = data_for_SingleR, ref = hpca.se, labels = hpca.se$label.main) #
data.hesc
```

```{r}
# table of seurat and singleR
table(data.hesc$labels,meta$seurat_clusters)
```
```{r}
# plot UMAP
data@meta.data$labels <- data.hesc$labels
DimPlot(data, group.by = c('seurat_clusters', 'labels'), reduction = 'umap')
```
The effect of annotation seems not good.

```{r}
data_1 <- data
data_1.hesc <- SingleR(test = data_for_SingleR, ref = list(BP = bpe.se, HPCA = hpca.se),
                       labels = list(bpe.se$label.main, hpca.se$label.main))
table(data_1.hesc$labels, meta$seurat_clusters)
```
```{r}
data_1@meta.data$lables <- data_1.hesc$labels
DimPlot(data_1, group.by = c('seurat_clusters', 'labels'), reduction = 'umap')
```

```{r}
### Annotation diagnosis
plotScoreHeatmap(data.hesc)
```

```{r}
# Plot the distribution of deltas 
# (i.e., the gap between the assignment score for the assigned label 
# and those of the remaining labels) across cells assigned to each reference label.
plotDeltaDistribution(data.hesc, ncol = 3)
```

```{r}
# compare with cluster results
tab <- table(label = data.hesc$labels,
             cluster = meta$seurat_clusters)
pheatmap(log10(tab + 10))
```

```{r}
data <- readRDS("RData/E6_bin200_seurat_result.rds")
data
meta <- data@meta.data
meta
```


```{r} 
ref = readRDS("Data/5patients_nG3000_PC30_Res0.4_Celltype.rds")
ref
```

```{r}
cell_type <- ref@meta.data %>% select(celltype.singResource)
cell_type
```

```{r}
ref.se <- SummarizedExperiment(assays=list(counts=ref[['RNA']]@counts), colData = cell_type) %>% logNormCounts()
ref.se
```
```{r}
saveRDS(ref.se, file = "RData/ref_se.rds")
```

```{r}
data_for_SingleR <- GetAssayData(data, slot="data")
```

```{r}
data.hesc <- SingleR(test = data_for_SingleR, ref = ref.se, labels = ref.se$celltype.singResource)
```

```{r}
data.hesc
```
```{r}
table(data.hesc$labels,meta$seurat_clusters)
```

```{r}
data@meta.data$labels <- data.hesc$labels
DimPlot(data, group.by = c('seurat_clusters', 'labels'), reduction = 'umap')
DimPlot(data, reduction = "umap", label = TRUE)
DimPlot(data, group.by = 'labels', reduction = 'umap')
```

```{r}
output_path = 'Output/'
write.csv(data.hesc, paste(output_path, 'SingleR_results.csv', sep = ""))
```

```{r}
### Annotation diagnosis
plotScoreHeatmap(data.hesc)
```
Ideally, each column of the heatmap should have one score that is obviously larger than the rest, indicating that it is unambiguously assigned to a single label.

```{r}
# Plot the distribution of deltas 
# (i.e., the gap between the assignment score for the assigned label 
# and those of the remaining labels) across cells assigned to each reference label.
plotDeltaDistribution(data.hesc, ncol = 3)
```

```{r}
# compare with cluster results
tab <- table(label = data.hesc$labels,
             cluster = meta$seurat_clusters)
pheatmap(log10(tab + 10))

```
SingleR在本地运行的结果与在与平台上运行的结果不同

自动注释的软件对于我们的研究工作能够起到一定的先导意义，不过这种先导意义一定是建立在对分析过程的理解之上的

```{r}
pruned.labels <- data.hesc$pruned.labels
length(pruned.labels)
```

```{r}
data@meta.data$pruned.labels <- pruned.labels
data
```

```{r}


plot <- 
  ggplot(data@meta.data, aes(x = x, y = y, color = factor(pruned.labels)))+
  geom_point() +
  theme_minimal()
print(plot)

```

cell type with their spatial location (with to find some patterns in Tumor)

the results in Cloud in much better (do not know why)

